Online Social Cohesion Reflects Real-World Group Action in Syria during the Arab Spring

### 2020-05-31 manuscript.Rmd
add literature review section

abstract: |
  Social mobilization requires effective collective coordination. In recent 
  years, individuals involved in social movements have taken to social media
  platforms to reach broader audiences at rapid speeds. Effective virtual 
  mobilization similarly requires social cohesion and coordination, yet the 
  degree to which virtual mobilization reflects or drives real-world movements
  is unclear. Here, we explore the dynamics of real-world events and social 
  cohesion on Twitter in Syria during the Arab Spring. Using the nonlinear 
  methods cross-recurrence quantification analysis and windowed cross-recurrence
  quantification analysis, we investigate if counts of all daily events, daily 
  positive events, and daily negative events are coupled with social cohesion 
  found in Syrian tweets. Results indicate that virtual social cohesion is 
  coupled with the counts of all and negative events each day but not with 
  counts of positive events. This suggests that virtual mobilization does reflect 
  real-world movements but only for salient, frequent and negative events. We 
  conclude with a discussion of implications and applications of nonlinear 
  methods in political science research.
  <!-- https://tinyurl.com/ybremelq -->

keywords          : "social cohesion, recurrence quantification analysis, complex systems, Twitter, Arab Spring"
wordcount         : "X"

@@

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
appendix          :

@@

# turn on caching
knitr::opts_chunk$set(cache=TRUE, autodep=TRUE, cache.lazy=FALSE)

@@

## Arab Spring
In December 2010, a Tunisian man named Mohamed Bouazizi was harassed and
unjustifiably stripped of his vendor cart by municipal officials. His
humiliation led him to set himself on fire in protest to the authoritarian
rule, ultimately leading to his death several days later. His actions served
as a catalyst for the Tunisian Revolution, and one week after his death,
Zine El Abidine Ben Ali---the longstanding authoritarian president---fled the
country. The events of Tunisia rapidly spread to surrounding countries 
also subject to unjust authoritarian rule, including Yemen, Egypt, Libya,
and Syria.

These mass movements would not have been possible without
efficient large-scale coordination, both physically and virtually on social
media platforms. Along with Facebook, Twitter has been credited as essential 
to the Arab Spring---both for protesters (as a means to coordinate with one 
another and to communicate with the world) and for the government response
(to monitor and track protestors to crack down on dissent) [@comninos2011twitter]. 
The events of Arab Spring have been
thoroughly discussed within the political science literature
[e.g., @lutterbeck2013arab; @koehler2017political], but to our knowledge,
the direct coupling between Syrian Twitter use and international action
has not been investigated.

## Social Cohesion and Social Mobilization
Real-world collective action is known to bring about a sense of togetherness,
also known as “collective effervescence” [@durkheim1995elementary]. This
feeling of being a part of something larger than oneself leads individuals
to act in ways that would not have on their own. Being overcome with this
feeling can lead to violence [@collins2009violence; @spaaij2014sports].
Conversely, increased exposure
to violence can result in positive social cohesion. @gilligan2014civil
found that exposure to fatal wartime violence resulted in increased prosocial
motivations, including cooperative behaviors, feelings of trust, and overall
community-level social cohesion.

Just as real-world social cohesion can be
quantified, so can online cohesion. While this can be quantified in many ways,
one kind of cohesion can be found in the similarity in language use on social media.
This implicit or explicit similarity in language use can indicate shared
beliefs, thoughts, and feelings, as has been studied in research on language style
in psycholinguistics and social psychology 
[@gonzales2010language; @tausczik2010psychological; @brennan1996conceptual]; 
although previous work on language style
has often focused on individuals and small groups, this can be scaled up to examine
massive social-level changes [@michel2011quantitative; @fusaroli2015timescales]. 
This kind of massive online social cohesion can lead to real-world
social change.

Combining these separate threads, it is reasonable to expect that real-world actions 
and online social interactions would be enmeshed, driving and being driven by 
one another. Rather than expecting that only real-world social cohesion would be
driven by real-world action, we would expect interaction between real-world and
virtual dynamics: In this case, exposure to real-world violence could result in social
cohesion in the virtual world. Increased social cohesion is thus potentially
an indicator and a result of increased social organization and mobilization 
on the ground.

Social media platforms provide one way of enabling this kind of online social
cohesion by offering distributed information-gathering and real-time information
dissemination. Given its accessibility and proficiency as an information
exchange tool, Twitter has provided individuals involved in social movements
an opportunity to complement or even replace real-world social mobilization.
For example, during the 2012 Italian protests for global economic change, the riots in
Rome were mirrored by Twitter activity both on the day of the protests and in the
days following them [@vicari2013public]. Tweeting was
determined to be more effective locally in discussing real-world, real-time
events than traditional mainstream media platforms. As more and more
individuals rely on social media to obtain local news and information, it
becomes increasingly important to understand just how these micro-blogs
reflect real-world action.

## Using Nonlinear Methods to Capture Complex Relationships
Given the hypothesized interrelationship between the real-world
and online actions---and the notorious messiness of naturally occurring
data---it is critical to find appropriate analysis tools that can
identify the complex interconnections.
Traditional linear analyses have led to many notable discoveries in the
domains of psychology and political science, but time series data pose
problems that are not easily overcome with these methods. Events have
historically been conceptualized as linear chains of smaller, sequential
events with one occurrence leading to the next. 
Certain kinds of data are relatively well-suited to these analyses,
but messy, complex, variable data from real-world events can pose
challenges from overpowered samples to violation of underlying
statistical assumptions [@paxton2017finding].

At the same time, the foundational theoretical assumptions of a 
linear world has been
increasingly replaced by ideas from dynamical systems theory. Rather than
a chain of events, time series data can be thought of as nested occurrences,
where mutually constrained parts (e.g., citizens, governments, real-world
events, social media platforms) are constantly interacting, leading to
more global behaviors that are larger than the sum of the parts
[@turvey2018lectures]. Linear analyses flatten these nonlinear patterns in 
time series data and ignore the variability
that is fundamental to these systems and their evolution over time. 
However, when analyzed as dynamical systems, many phenomena
exhibit new nonlinear patterns that were formerly overlooked. 
Nonlinear methods provide means for capturing the rich temporal
dynamics and variability that unfold in time series data.

Although a number of other nonlinear methods exist, 
the primary nonlinear method used in the current work is cross-recurrence
quantification analysis [CRQA; @zbilut1998detecting] and its offshoot,
windowed cross-recurrence quantification analysis [WCRQA; @webber2005recurrence].
CRQA has successfully been used to gain a better understanding of many
complex human phenomena, including a variety of human social interactions
[for a review, see @fusaroli2014analyzing]. 
Much as it has for psychology, we argue that the use of nonlinear methods in
the investigation of political science phenomena will lead to new insights
and the development of more robust theories.

## The Present Work
Social media has become a primary platform for social movements and
activism across the world, but the extent to which virtual activism
reflects and drives real-world mobilization is still largely unknown. 
In the current work, we
conceptualize the physical and virtual movements as a dynamical system,
both constantly interacting and influencing the other. We aim to study
this by analyzing the real-world events and online social cohesion
during nearly three months of the Arab Spring. Both real- and
virtual-world collective coordination were powerful tools in the events
of the Arab Spring and are undeniably linked. This study aims to explore
the dynamics of Syrian Twitter users, who used the social media platform to raise
awareness about the Arab Spring and bring about mass mobilization.

Using cross-recurrence quantification analysis [CRQA; @zbilut1998detecting]
and windowed cross-recurrence quantification analysis
[WCRQA; @webber2005recurrence], we analyzed the co-evolution of social
cohesion online and real-world events. Specifically, using heterogeneous
data from March 30, 2012, to June 15, 2012, we compared the social cohesion
of Syrian tweets with the counts
of all, positive, and negative real-world events derived from gold-standard
international event data [i.e., the Integrated Crisis Early Warning
System; @DVN/28075_2015].

We hypothesized that the count of all events and the count of negative
events would show high levels of determinism, as this time period was
marked by daily violence which often leads to increased salience. We also
hypothesized that the count of positive events would not show patterns of
determinism, as these events were likely overshadowed by the salient
negative daily occurrences. 
After testing these hypotheses, we then conducted exploratory analyses 
to further examine the patterns identified with CRQA. Using WCRQA,
we sought to identify shifts in the
dynamics of the systems that could then be linked to real-world shifts in
conflict frequency.
Through these analyses, our two primary goals in the current work were to uncover
(1) how different intensities during times of turmoil may enhance or deter
social cohesion in social media communication and (2) how social mobilization
may be evident in the dynamic coupling of cohesion and real-world events.

An additional focus of the current work is to introduce cross-recurrence
quantification analysis---a longstanding method from physics that has 
since become influential in a variety of other fields---to political 
science. We do so by providing detailed descriptions of the methods and
providing clear directions about interpretation and significance testing
even with relatively small-*n* or case studies. After demonstrating these
methods' utility with this dataset, we close the paper with specific
suggestions to political scientists interested in incorporating these
methods into their work.

# Method
## Materials
### Tweet Corpus

```{r tweet-count, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, results='hide'}

# read in social cohesion data
cohesion_df <- read.csv("./data/raw/Syria-social_cohesion.csv")

The corpus of tweets used included `r sum(cohesion_df$n)` Syrian tweets
occurring between March 30--June 15, 2012. Using the Twitter
streaming API, servers at the University of Texas at Austin collected
the tweets. The servers were run on a stable (LTS) version of Ubuntu
Linus (v. 10.04) and were connected directly to the high-bandwidth UT network
maintained by IT services.

Tweets collected contained one or more of the following keywords:
“syria,” “syrian,” “damascus,” “homs,” “al-assad,” and “sunni.” These
keywords were chosen based on the most distinctive terms used in news
articles covering the Arab Spring in Syria at the time. A node.js script
was used to reinitialize the connection to Twitter's servers whenever the
script timed out due to a period of silence. The script requested data from
the API endpoint using basic HTTP authentication (OAuth was optional, at
the time) at https://stream.twitter.com/1.1/statuses/filter.json, and stored
the data directly to the local hard disk. The tweets were then post-processed
with a Python script to convert the JSON-formatted tweets to tab-separated
files, which were then compressed to maximize storage efficiency.

All scripts used for scrapping tweets and compressing data can be found at https://github.com/chbrown/twilight.

### Event Data
Event data were obtained from the 2012 Integrated Crisis Early Warning
System [ICEWS; @DVN/28075_2015]. The event data were first filtered to
remove incomplete and incorrectly formatted data. They were then filtered
to include only events that were directed at (i.e., `target`) Syria ($n = 6300$).
To explore the impact of outwardly directed Syrian events, the data were
filtered again to include events that were both directed at (i.e., `target`)
or directed by (i.e., `source`) Syria ($n = 7990$).

## Data Preparation
### Social Cohesion Metric
The social cohesion or linguistic similarity of daily tweets was calculated
by quantifying the frequency of shared content words with respect to those
words’ frequencies in the entire tweet corpus. If a linguistic term has a
high frequency in one tweet but has a low frequency in the whole tweet
corpus, then it is more likely a content word and is thus given more weight
in calculating social cohesion. 
This method is similar to inverse document frequency (IDF).
For more information on IDF, see @jones1972statistical and
@robertson2004understanding.

The tweets were first sorted into ascending order by timestamp. They were
then grouped into successive sets of five tweets. Inverse frequency
weighting for the overlap words and all words was conducted for each pair
of two tweets in the group of five, then divided to get the social cohesion
score for that pair (see Eq. 1-3). This resulted in ten social cohesion
values for each group of five tweets; these values were then averaged to get
a social cohesion score for the entire set of five tweets. Finally,
the mean of all average cohesion values for all groups was taken to get
a social cohesion score for each day.

\begin{align}
\text{Inverse frequency weighting of overlap words = }\sum_{i=1}^k \frac{log(f(t_i, d1) + f(t_i, d2) + 1)}{log(f(t_i) + 1)}
\end{align}

\begin{align}
\text{Inverse frequency weighting of all words = }\sum_{i=1}^n \frac{log(f(t_i, d1) + f(t_i, d2) + 1)}{log(f(t_i) + 1)}
\end{align}

\begin{align}
\text{Two tweets similarity =}\frac{\text{inverse frequency weighting of overlap words}}{\text{inverse frequency weighting of all words}}
\end{align}

The social cohesion calculation was performed on the tweet corpus
containing (recalculate with new date span) English and (recalculate with
new date span) Arabic tweets. All hashtags and URLs were included in the
analysis.

### ICEWS Event Time Series
Using the Integrated Crisis Early Warning System (ICEWS) developed at the
Defense Advanced Research Project Agency, the daily events were
scored with an intensity rating ranging from -10 to 10. Negative 
events---such as riots and attacks---are coded with low negative numbers, while
positive or peaceful events are coded with high positive numbers. This
well-validated coding scheme allows for numerical quantification of
different event types.

For both the target only and the target and source data described above, time series of
the count of all events were generated by summing the total number of events
on each day, irrespective of ICEWS score. Similarly, time series of count
of positive events were generated by summing the number of ICEWS events
for each day that had an intensity rating greater than zero. Lastly,
time series of count of negative events were generated by summing the number
of ICEWS events for each day that had an intensity rating less than zero.

## Analyses
### Cross-Recurrence Quantification Analysis
We first used cross-recurrence quantification analysis
[CRQA; @zbilut1998detecting] to investigate the temporal patterns of
social cohesion and real-world events in Syria during Arab Spring. CRQA
is an extension of RQA, which captures the patterns of revisited states of
a single time series [@zbilut1992embeddings; @webber1994dynamical]. CRQA
allows for two time series to be projected onto one another to investigate
shared states and trajectories.
In CRQA, we track these shared states by identifying *recurrent points*,
or intersections of time during which two signals shared or revisited the same
state. These shared states can be at the same time or at later points in
the time series. We can further chart shared trajectories by examining
the occurrences and patterns of multiple sequential recurrent points.

To perform CRQA, the time series must be converted to the same
scale, while still preserving a sampling rate that respects the native
timescale of the system
[cf. Nyquist frequency in signal processing; @grenander1959probability].
Because the social cohesion and event data have different scales and values,
we converted each time series (e.g., Figure \@ref(fig:raw-ts)) into deciles
(e.g., Figure \@ref(fig:deciled-ts)) to analyze the relative dynamics of different
levels of social cohesion and counts of all, positive, and negative events
on each day. 
We implemented CRQA using the `crqa` package [@R-crqa] for R [@R-base].

#### Visualization
A cross-recurrence plot (CRP) is the visual representation of CRQA. To construct
a cross-recurrence plot, two time series are plotted against each other on the
$x$ and $y$ axes. At every point where they share the same state, a point
in placed on the plot to indicate recurrence. (For categorical CRQA, this process
is similar to the Cartesian product.)

We can visually analyze these plots
to reveal patterns of structure and recurrence. The line $y = x$ is known
as the *line of identity*. This line falls where the time series match
temporally. A filled line would indicate perfect in-phase synchrony between the
two time series. See
Figures \@ref(fig:plot-rp-targ-all)-\@ref(fig:plot-rp-source-targ-neg)
for examples of RPs.

#### Metrics
While visual inspections can be informative, we can derive a variety of 
quantitative metrics from the cross-recurrence plot. Each of these metrics
provides information into the pattern and dynamics of the new shared system
between the original time series. These metrics are also critical for inferential
statistics. We will focus most directly on two of these metrics: 
*recurrence rate* and *determinism*.

The percentage of recurrent points to total possible points is a metric
called the *recurrence rate* (*RR*). This metrics gives a general description of
how frequently the time series are sharing the same states. In the current
study, a recurrent point would indicate that the level of social cohesion
and frequency of events were at the same intensity.

Recurrent points that occur in successive moments form line structures.
These are conceptualized as shared trajectories, where the two time series
are moving together across time in a shared state. The percentage of
recurrent points on the plot that fall along diagonal line structures 
(i.e., two or more
consecutive points) is a metric called *percent determinism* 
(or simply *determinism*; *DET*). For the
current project, a high DET would indicate that not only are the levels of
social cohesion and frequency of events occurring at similar intensities,
but they are doing so across consecutive days.

A number of other metrics provide insight not only into the amount of structure
between the systems but about the nature of this shared structure.
The average length of the line structures is called *mean line* (L), which can
be conceptualized as the average amount of time that the two time series move
together before one or both shift states. The longest line length is called
*max line* (*maxL*); it indicates the longest instance of being “stuck” in a
particular state together. The total *number of lines* on the plot (*NRLINE*) 
scales with structure: The more lines that exist, the more coupled the systems are. The
complexity or variability of line lengths is captured by a metric called
*entropy* (*ENTR*)---which may also be normalized by the total number of lines
(i.e., *normalized entropy*; *rENTR*)---and indicates
the stability and predictability of the system. When one of the signals gets
“stuck” in a state and the other moves on to other states, this results in
vertical line structures, which are quantified in *laminarity* (*LAM*;
i.e., the proportion of points forming vertical line structures) 
and trapping time (*TT*; i.e., the average length of the
vertical trajectories). For more detail on CRQA metrics, see @R-crqa.

#### Parameters
To conduct the analyses on the decile data, the parameters were set to those
typically used in categorical CRQA: Delay was set to $0$, embedding dimension
was set to $1$, and radius was set to $.001$ to allow for only exact recurrent
matches. No normalization was done, and lines were considered two or more
consecutive recurrent points. The Theiler window was set to $0$ to include
points along the line of identity. For more details on the CRQA parameters and
how they impact the analysis, see @R-crqa.

#### Permutation Test
Unlike some nonlinear analyses (e.g., fractal analyses), CRQA metrics do not have
inherently meaningful values. As a result, CRQA is best used as a relative
metric---that is, comparing the CRQA values across conditions or subsets
of the data. Typically, meaningful differences between values are determined with
traditional inferential statistics. While this may not be a problem for 
experimentally derived datasets, it can pose a barrier to researchers using
small-$n$ data or conducting data-informed case studies.

For CRQA, this can be addressed using *permutation tests*.
Similar to more established methods of statistical significance in CRQA research
[which create surrogate time series; e.g., @paxton2017interpersonal],
the permutation test approach to significance testing allows the researcher
to use a time series as its own baseline by testing whether the *structure*
of the unfolding events of the time series---rather than the *raw frequency* of the 
events within the time series---cohere together more than would be expected
by chance. A permutation test is similar to
bootstrapping, except that the sampling of observations is done without
replacement. In this approach, we simply create a large number of permutations
of the original time series (i.e., breaking up the dependency across time points but
preserving the frequencies of the events), conduct CRQA for each of those
permutations, and then examine whether the real time series' CRQA metrics
exceed those that we would expect to see by chance. The proportion of times that
the real time series' values exceeds the surrogate time series' values can
be interpreted as the alpha criterion to determine significance. 
For more information on permutation
tests, see @good2013permutation.

Using the real deciles of social cohesion, count of all events,
count of positive events, and count of negative events, we conducted a series
of permutation tests 
(one for the comparison between social cohesion and each type of event count)
to determine statistical difference from chance for each CRQA metric.
Using the `sample` function in base R [@R-base], $1000$ permuted time series were
created for each of the real time series.  
CRQA was then run $1000$ times for social cohesion and count of all
events, social cohesion and count of positive events, and social cohesion and
count of negative events using the permuted time series. This resulted in
$1000$ results for each of the CRQA metrics for each combination of time series.
Statistical significance from chance was set at the 95th and 99th percentiles
of the metrics from the permuted CRQ analyses. Given that a permutation test
is similar to bootstrapping but does not include replacement, all points were
included in each CRQA analysis, leaving RR constant across all permutations. Thus,
RR is not included in the inferential tests.

### Windowed Cross-Recurrence Quantification Analysis
We used windowed cross-recurrence quantification analysis
[WCRQA; @webber2005recurrence] to further investigate the evolution of the
temporal dynamics of social cohesion and real-world events. In this analysis,
a smaller “window” of time points is used to calculate CRQA and capture more
fine-grained dynamics of two time series. The window remains the same size and
“slides” diagonally up the line of identity. The changes in the CRQA metrics
across the windows can reveal changes in dynamics that can then be linked to
external, real-world events. 
We implemented WCRQA using the `crqa` package [@R-crqa] for R [@R-base].

#### Metrics
Conducting WCRQA results in CRQA metrics (e.g., RR, DET, ENTR) for each
respective window. These metrics can then be plotted as a time series to
visualize their evolution over time. For these analyses, we decided to
focus on recurrence rate and determinism.

#### Parameters
In addition to the CRQA parameters, two other parameters are required when
conducting WCRQA. First, the *window size*
will determine how many observations are used in each CRQA analysis. We
conducted WCRQA with a window size of $14$ to investigate the dynamics over
two weeks and allow for the capture of more structure and thus a more
meaningful interpretation of DET.

The second parameter to be chosen is a *window step size*. 
Window step size is how far the window moves up the line of identity
before conducting CRQA again. We chose $1$ for the step size to understand
how the metrics evolve from day to day.

All other CRQA metrics remained the same as those previously noted.

#### Visualization
To visualize WCRQA, the resulting metrics from each window can be plotted
as a time series. This allows for visual interpretation of the pattern or
fluctuation of the metrics across time. These dynamics can then be linked
to other external events. See Figures \@ref(fig:plot-wcrqa-targ-all) and
\@ref(fig:plot-wcrqa-source-targ-all) for examples.

>>>>

abstract: |
  Social mobilization requires effective collective coordination. In recent years, individuals involved in social movements have taken to social media platforms such as Twitter to reach broader audiences at rapid speeds. Effective virtual mobilization similarly requires social cohesion and coordination, yet it is still unclear if virtual mobilization reflects real-world movements. Here, we explore the dynamics of real-world events and social cohesion on Twitter in Syria during the Arab Spring. Using the nonlinear methods cross-recurrence quantification analysis and windowed cross-recurrence quantification analysis, we investigate if counts of all daily events, daily positive events, and daily negative events are coupled with social cohesion found in Syrian tweets. Results indicate that virtual social cohesion is coupled with the counts of all and negative events each day, but not with counts of positive events. This suggests that virtual mobilization does reflect real-world movements, but only for salient, frequent and negative events. We conclude with a discussion of implications and applications of nonlinear methods in political science research.
  <!-- https://tinyurl.com/ybremelq -->

keywords          : "Social cohesion, recurrence quantification analysis, complex systems, Twitter, nonlinear dynamics, Arab Spring"
wordcount         : "X"

@@

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
appendix          : "appendix.Rmd"

@@

# turn on caching
knitr::opts_chunk$set(cache=TRUE, autodep=TRUE, cache.lazy=FALSE)
#knitr::opts_chunk$set(fig.pos = "!H", out.extra = "")

@@

## Arab Spring
In December 2010, a Tunisian man named Mohamed Bouazizi was harassed and unjustifiably stripped of his vendor cart by municipal officials. His humiliation led him to set himself on fire in protest to the authoritarian rule, ultimately leading to his death several days later. His actions served as a catalyst for the Tunisian Revolution, and one week after his death, the long-withstanding authoritarian president, Zine El Abidine Ben Ali, fled the country. The events of Tunisia rapidly spread to surrounding countries that were also subject to unjust authoritarian rule, such as Yemen, Egypt, Libya, and Syria. These mass movements would not have been possible without efficient interpersonal coordination, both physically and virtually on social media platforms, such as Twitter. The events of Arab Spring have been thoroughly discussed within the political science literature [e.g., @lutterbeck2013arab; @koehler2017political] but to our knowledge, the use of Twitter by Syrian citizens has not been investigated.

## Literature review

## Social Cohesion
Real-world collective action is known to bring about a sense of togetherness, also known as “collective effervescence” [@durkheim1995elementary]. This feeling of being a part of something larger than oneself leads individuals to act in ways that would not have on their own. Being overcome with this feeling is known to lead to violence [@collins2009violence; @spaaij2014sports]. While feelings of social connection may lead to violence, increased exposure to violence can result in feelings of social cohesion. @gilligan2014civil found that exposure to fatal wartime violence resulted in increased prosocial motivations, including cooperative behaviors, feelings of trust, and overall community-level social cohesion. Just as real-world social cohesion can be quantified, the same is true for online cohesion. Strategic communication on social media in the form of social cohesion of language can lead to real-world social change, and exposure to real-world violence could result in social cohesion in the virtual world. Increased social cohesion is thus potentially an indicator of increased social organization and mobilization on the ground.

## Social Media as a Platform for Mobilization
Social media platforms, such as Twitter, offer rapid real-time dissemination of information. Given its accessibility and proficiency as an information exchange tool, individuals involved in social movements have taken to them in conjunction, or in replacement, of real-world social mobilization. During the 2012 Italian protests for global economic change, the riots in Rome were mirrored by Twitter activity both on the same day and in the following days after the protests [@vicari2013public]. Tweeting was determined to be more effective locally in discussing real-world, real-time events than traditional mainstream media platforms. As more and more individuals rely on social media to obtain local news and information, it becomes increasingly important to understand just how these micro-blogs reflect real-world action.

## Use of Nonlinear Methods in Naturally Occurring Data
Traditional linear analyses have led to many notable discoveries in the domains of psychology and political science, but continuous data pose problems that are not easily overcome with these methods. Events have historically been conceptualized as linear chains of smaller, sequential events with one occurrence leading to the next. This notion has been increasingly replaced by ideas from dynamical systems theory. Rather than a chain of events, continuous data can be thought of as nested occurrences, where mutually constrained parts (e.g., citizens, governments, real-world events, social media platforms) are constantly interacting, leading to more global behaviors that are larger than the sum of the parts [@turvey2018lectures]. When analyzed as dynamical systems, many phenomena exhibit new nonlinear patterns that were formerly overlooked. Continuous data subject to linear analyses becomes flattened, and the variability that is fundamental to these systems and their evolution over time is ignored. Nonlinear methods provide means for capturing the rich temporal dynamics and variability that unfold in continuous data.
The primary methods used in the current work is cross-recurrence quantification analysis [CRQA; @zbilut1998detecting] and it’s off-shoot, windowed cross-recurrence quantification analysis [WCRQA; @webber2005recurrence]. CRQA has successfully been used to gain a better understanding of many complex human phenomena, such as human social interactions [for a review, see @fusaroli2014analyzing] and interpersonal synchrony [@paxton2017interpersonal]. We argue that the use of nonlinear methods in the investigation of political science phenomena will lead to new insights and development of more robust theories.

## The Present Work
Social media has become a primary platform for social movements and activism across the world, but the extent to which virtual activism reflects real-world mobilization is still largely unknown. We conceptualize the physical and virtual movements as a dynamical system, both constantly interacting and influencing the other. Both real- and virtual-world collective coordination were powerful tools in the events of Arab Spring, and are undeniably linked. This study aims to explore the dynamics of Syrian individuals who took to Twitter to raise awareness about the events of Arab Spring and bring about mass mobilization. Using cross-recurrence quantification analysis [CRQA; @zbilut1998detecting] and windowed cross-recurrence quantification analysis [WCRQA; @webber2005recurrence], we analyzed the dynamics of social cohesion of tweets from March 30, 2012 to June 15, 2012 and the counts of all, positive, and negative real-world events on those respective days. We hypothesized that the count of all events and the count of negative events would show high levels of determinism, as this time period was marked by daily violence which often leads to increased salience. We also hypothesized that the count of positive events would not show patterns of determinism, as these events were likely overshadowed by the salient negative daily occurrences. Using WCRQA, we hoped to identify shifts in the dynamics of the systems that could then be linked to real world shifts in conflict frequency. As social cohesion has never been analyzed using nonlinear methods, the analyses contained herein are largely exploratory. The goal of this endeavor is to gain a better understanding of how events of different intensities during times of turmoil may enhance or deter social cohesion in social media communication, and how social mobilization may be evident in the dynamic coupling of cohesion and real-world events.

# Method
## Materials
### Tweet Corpus
The corpus of tweets used included 8,084,644 Syrian tweets occurring between March 1, 2012 and June 15, 2012. A server malfunction during data collection resulted in missing data from March 12, 2012 to March 29, 2012. For the current analyses, the data were truncated to remove missing data, making the final dataset from March 30, 2012 to June 15, 2012 (n = 78). The tweet corpus was scraped from Twitter by [insert search parameters here].

### Event Data
Event data were obtained from the 2012 Integrated Crisis Early Warning System [ICEWS; @DVN/28075_2015]. The event data were first filtered to remove incomplete and incorrectly formatted data. They were then filtered to include only events that were directed (i.e., target) at Syria (n = 6300). To explore the impact of outwardly directed Syrian events, the data were filtered again to include events that were both directed at (i.e., target) or directed by (i.e., source) Syria (n = 7990).

## Data Preparation
### Social Cohesion Metric
The social cohesion or linguistic similarity of daily tweets was calculated by quantifying the frequency of shared content words with respect to those words’ frequencies in the entire tweet corpus. If a linguistic term has a high frequency in one tweet but has a low frequency in the whole tweet corpus, then it is more likely a content word and is this given more weight in calculating social cohesion.
The tweets were first sorted into ascending order by timestamp. They were then grouped into successive sets of five tweets. Inverse frequency weighting for the overlap words and all words was conducted for each pair of two tweets in the group of five, then divided to get the social cohesion score for that pair (see Eq. 1-3). This resulted in ten social cohesion values for each group of five tweets, which were then averaged. Finally, the mean of all average cohesion values for all groups was taken to get a social cohesion score for each day.

(insert social cohesion equations here)

The social cohesion calculation was performed on a subset of the corpus containing (recalculate with new date span) English and (recalculate with new date span) Arabic tweets. All hashtags and URLs were included in the analysis. This method is similar to that of inverse document frequency (IDF). For more information on IDF, see @jones1972statistical and @robertson2004understanding.

### ICEWS Event Time Series
Using the Integrated Crisis Early Warning System (ICEWS) developed at the Defense Advanced Research Project Agency, the daily events in Syria were scored with an intensity rating ranging from -10 to 10. Negative events, such as riots and attacks, are coded with low negative numbers while positive, peaceful events are coded with high positive numbers. This well-validated coding scheme allows for numerical quantification of different event types.
For both the target only and the target and source data, time series of the count of all events were generated by summing the total number of events on each day, irrespective of ICEWS score. Similarly, time series of count of positive events were generated by summing the number of ICEWS events for each day that had an intensity rating greater than zero. Lastly, time series of count of negative events were generated by summing the number of ICEWS events for each day that had an intensity rating less than zero.

## Analyses
### Cross-Recurrence Quantification Analysis
We first used cross-recurrence quantification analysis [CRQA; @zbilut1998detecting] to investigate the temporal patterns of social cohesion and real-world events in Syria during Arab Spring. As an extension of RQA, which captures the patterns of revisited states of a single time series [@zbilut1992embeddings; @webber1994dynamical], CRQA allows for two time series to be projected onto one another to investigate shared states and trajectories. Nonlinear analyses such as CRQA can reveal dynamics that are not captured by traditional linear methods. In order to perform CRQA, the time series need to be converted to the same scale, while still preserving a sampling rate that respects the native timescale of the system [cf. Nyquist frequency in signal processing; @grenander1959probability]. We converted each time series (e.g., Figure 1) into deciles (e.g., Figure 2) to analyze the dynamics of different levels of social cohesion and counts of all, positive, and negative events on each day. In CRQA, a recurrent point is when two signals share the same state. These shared states can be at the same time or at later points in the time series.

#### Visualization
A recurrence plot (RP) is the visual representation of CRQA. To construct a recurrence plot, two time series are plotted against each other on the X and Y axes, and at every point where they share the same state, and point in placed on the plot to indicate recurrence. Visually analyzing these plots can help reveal patterns of structure and recurrence. The line y = x is known as the line of identity. This line falls where the time series match temporally, and a filled line would indicate perfect synchrony between the two time series. See Figures 3-5 for examples of RPs.

#### Metrics
The percentage of recurrent points to total possible points is a metric called recurrence rate (RR). This metrics gives a general description of how frequently the time series are sharing the same states. In the current study, a recurrent point would indicate that the level of social cohesion and frequency of events were at the same intensity.
Recurrent points that occur in successive moments form line structures. These are conceptualized as shared trajectories, where the two time series are moving together across time in a shared state. The percentage of recurrent points that make up diagonal line structures (i.e., two or more consecutive points) is a metric called percent determinism (DET). For the current project, a high DET would indicate that not only are the levels of social cohesion and frequency of events occurring at similar intensities, but they are doing so across consecutive days.
The average length of the line structures is called mean line (L). This can be conceptualized as the average amount of time that the two time series move together before one or both shift states. The longest line length is called max line (maxL). maxL indicates the longest instance of being “stuck” in a particular state together. The total number of lines is called NRLINE. The more lines there are, the more coordinated the system may be. The complexity or variability of line lengths is captured by a metric called entropy (ENTR) and entropy while accounting for the total number of lines in the plot is normalized entropy (rENTR). These entropy metrics help indicate the stability and predictability of the system. When one of the signals gets “stuck” in a state and the other moves on to other states, this results in vertical line structures. The proportion of points forming vertical line structures is a metric called laminarity (LAM) and the average length of the vertical trajectories is called trapping time (TT). For more detail on CRQA metrics, see @R-crqa.

#### Parameters
To conduct the analyses on the decile data, the parameters were set to those typically used in categorical CRQA. Delay was set to zero, embedding dimension was set to 1, and radius was set to .001 to allow for only exact recurrent matches. No normalization was done, and lines were considered two or more consecutive recurrent points. The Theiler window was set to zero to include points along the line of identity. For more details on the CRQA parameters and how they impact the analysis, see @R-crqa.

#### Permutation Test
Using the deciled, real time series of social cohesion, count of all events, count of positive events, and count of negative events, a permutation test was done to determine statistical difference from chance for each CRQA metric. Using the sample function in base R [@R-base], 1000 permuted time series were created for each of the real time series. A permutation test is similar to bootstrapping, except that the sampling of observations is done without replacement. CRQA was then run 1000 times for social cohesion and count of all events, social cohesion and count of positive events, and social cohesion and count of negative events using the permuted time series. This resulted in 1000 results for each of the CRQA metrics for each combination of time series. Statistical significance from chance was set at the 95th and 99th percentiles of the metrics from the permuted CRQ analyses. Given that a permutation test is similar to bootstrapping but does not include replacement, all points were included in each CRQA analysis, leaving RR the same across all analyses. Thus, RR is not included in the inferential tests. For more information on permutation tests, see @good2013permutation.

### Windowed Cross-Recurrence Quantification Analysis
We used windowed cross-recurrence quantification analysis [WCRQA; @webber2005recurrence] to further investigate the evolution of the temporal dynamics of social cohesion and real-world events. In this analysis, a smaller “window” of time points is used to calculate CRQA and capture more fine-grain dynamics of two time series. The window remains the same size and “slides” diagonally up the line of identity. The changes in the CRQA metrics across the windows can reveal changes in dynamics that can then be linked to external, real-world events.

#### Metrics
Conducting WCRQA results in CRQA metrics (e.g., RR, DET, ENTR) for each respective window. These metrics can then be plotted as a time series to visualize their evolution over time. For these analyses, we decided to visualize recurrence rate and determinism.

#### Parameters
In addition to the CRQA parameters, two other parameters are required when conducting WCRQA. First, a window size must be decided on. The window size will determine how many observations are used in each CRQA analysis. We conducted WCRQA with a window size of 14 to investigate the dynamics over two weeks and allow for the capture of more structure and thus a more meaningful interpretation of DET. The second parameter to be chosen is a step size. Window step size is how far the window moves up the line of identity before conducting CRQA again. We chose 1 for the step size to understand how the metrics evolve from day to day. All other CRQA metrics remained the same as previously noted.

#### Visualization
To visualize WCRQA, the resulting metrics from each window can be plotted as a time series. This allows for visual interpretation of the pattern or fluctuation of the metrics across time. These dynamics can then be linked to other external events. See Figures 9-14 and 16-21 for examples.

#### Permutation Test
To visualize the differences in the WCRQA metrics from those obtained by chance, a permutation test was done to set upper and lower bounds for the 95th and 99th percentiles. Using a sample size of 14, one-thousand permuted time series were generated without replacement for each real time series. The upper and lower 95th and 99th percentiles were set based on the distribution of CRQA metrics from these one-thousand CRQ analyses.
